---
output: github_document
---

[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/OscarGOGO/Makurhini?branch=master&svg=true)](https://ci.appveyor.com/project/OscarGOGO/Makurhini)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Makurhini

The goal of Makurhini is to provide a set of functions to estimate landscape fragmentation and connectivity metrics

## Installation

You can install the released version of Makurhini from [GitHub](https://github.com) with:

``` r
library(devtools)
install_github("OscarGOGO/Makurhini", dependencies = TRUE)
```

## Example
This is a basic example which shows you how to solve some common problems:

### Protected Connected Land (ProtConn)

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(Makurhini)
library(raster)
library(rgeos)
```

Protected areas:

```{r message=FALSE, warning=FALSE, echo=TRUE}
data("Protected_areas", package = "Makurhini")
data("regions", package = "Makurhini")
region <- regions[1,]
```


```{r cores, message=FALSE, warning=FALSE, echo=FALSE}
plot(Protected_areas, col="#beaed4", main="Protected areas",
     axes = TRUE, key.pos = 4)

plot(region, col="#7fc97f", main="Ecoregion",
     axes = TRUE, key.pos = 4)

```


```{r message=FALSE, warning=FALSE}
test <- MK_ProtConn(nodes = Protected_areas, region = region,
                    attribute = "Intersected area", area_unit = "ha",
                    distance = list(type= "centroid"),
                    distance_thresholds = 10000,
                    probability = 0.5, transboundary = 50000,
                    LA = NULL, plot = TRUE, dPC = FALSE,
                    write = NULL, SAGA = FALSE, intern = FALSE)
test$`Protected Connected (Viewer Panel)`
test$`ProtConn Plot`
```

###  Equivalent Connectivity (EC) 

Example with old-growth vegetation fragments of four times (T1, T2, T3 and T4).

```{r message=FALSE, warning=FALSE}
ruta <- system.file("extdata", "dECA_example.RData", package = "Makurhini")
load(ruta)

Max_attribute <- gArea(study_area) * 0.0001 #hectares
```

```{r message=FALSE, warning=FALSE}
dECA_test <- MK_dECA(nodes= forest_patches, attribute = NULL, area_unit = "ha",
                  distance = list(type= "centroid"), metric = "PC",
                  probability = 0.05, distance_thresholds = 5000,
                  LA = Max_attribute, plot= c("T1", "T2", "T3", "T4"))
dECA_test


```


### Citing Makurhini package
A formal paper detailing this packe is forthcoming, but until it is published, please use the something like the following to cite if you use it in your work:

<code> <i> Godínez-Gómez, O. and Correa Ayram C.A. 2020. Makurhini: An R package for analyzing landscape connectivity. https://github.com/OscarGOGO/Makurhini, DOI: 10.5281/zenodo.3748045 </i> </code>
