---
title: "Fragmentation statistics"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Fragmentation_statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
We explored the <strong> MK_Fragmentation() </strong> to estimate fragmentation statistics at the landscape and patch level.

```{r setup, message=FALSE, warning=FALSE}
library(Makurhini)
library(sf)
```

## Example database 
We loaded the "vegetation_patches" database with 142 patches of old-growth vegetation from the state of Chiapas, Mexico. The shapefile has 142 rows (patches), the first column corresponds to the id of the patch.

```{r message=FALSE, warning=FALSE}
data("vegetation_patches", package = "Makurhini")
nrow(vegetation_patches) # Number of patches
```

```{r someplot, message=FALSE, warning=FALSE, echo=FALSE}
library(mapview)
vegetation_patches$Legend <- "Old-growth vegetation fragments"
mapview(vegetation_patches, zcol ="Legend", col.regions= "#00B050")
vegetation_patches$Legend <- NULL
```

## MK_Fragmentation()

To define the edge of the patches we will use a distance of 500 m from the limit of the patches (Haddad et al. 2015).


```{r message=FALSE, warning=FALSE}
Fragmentation_test <- MK_Fragmentation(patches = vegetation_patches, edge_distance = 500, plot = TRUE,
                                       min_patch_area = 100, landscape_area = NULL, area_unit = "km2", perimeter_unit = "km")
```

### Exploring results

* The results are presented as a list, the first result is called *"Summary landscape metrics (Viewer Panel)"* and it has fragmentation statistics at landscape level.

```{r message=FALSE, warning=FALSE}
names(Fragmentation_test)
Fragmentation_test$`Summary landscape metrics (Viewer Panel)`
```

* The second output *"Patch statistics shapefile"* is a shapefile with patch level fragmentation statistics that can be saved using write_sf() from *'sf'* package (https://cran.r-project.org/web/packages/sf/index.html).

```{r message=FALSE, warning=FALSE}
head(Fragmentation_test[[2]])

#To save the shapefile 
#write_sf(Fragmentacion$`Patch statistics shapefile`, ".../folder/fragmentacion.shp")
```


### Viewing the results
We can visualize the static at the patch level using the default *plot()* function or other spatial information display packages like the *'tmap'* package, for example:

* Core area (%):
```{r message=FALSE, warning=FALSE}
library(tmap)
tmap_mode("plot")
tm_shape(Fragmentation_test[[2]]) + tm_fill("CAPercent", palette = c( "#762A83", "#AF8DC3", "#E7D4E8", "#D9F0D3", "#7FBF7B", "#1B7837"))+ tm_style("cobalt")+ tm_layout(legend.width = 0.43, legend.height = 0.43)
```

* Edge (%)

```{r message = FALSE, warning = FALSE}
tm_shape(Fragmentation_test[[2]]) + tm_fill("EdgePercent", palette = c("#542788", "#998EC3", "#D8DAEB", "#FEE0B6", "#F1A340", "#B35806"))+ tm_style("cobalt")+ tm_layout(legend.width = 0.43, legend.height = 0.43)
```

* Perimeter-area ratio (PARA)

```{r message = FALSE, warning = FALSE}
tm_shape(Fragmentation_test[[2]]) + tm_fill("PARA", palette = c("#542788", "#998EC3", "#D8DAEB", "#FEE0B6", "#F1A340", "#B35806"))+ tm_style("cobalt")+ tm_layout(legend.width = 0.43, legend.height = 0.43)
```

* Shape Index

```{r message = FALSE, warning = FALSE}
tm_shape(Fragmentation_test[[2]]) + tm_fill("ShapeIndex", palette = c("#542788", "#998EC3", "#D8DAEB", "#FEE0B6", "#F1A340", "#B35806"))+ tm_style("cobalt")+ tm_layout(legend.width = 0.43, legend.height = 0.43)

```

* Fractal Dimension Index

```{r message = FALSE, warning = FALSE}
tm_shape(Fragmentation_test[[2]]) + tm_fill("FRAC", palette = c("#542788", "#998EC3", "#D8DAEB", "#FEE0B6", "#F1A340", "#B35806"))+ tm_style("cobalt")+ tm_layout(legend.width = 0.43, legend.height = 0.43)
```

### Exploring other edge depths
We can make a loop where we explore different edge depths. In the following example, We will explore 10 edge depths (*edge_distance argument*): 100, 200, 300, 400, 500, 600, 700, 800, 900 and 1000 meters. We will apply the *'MK_Fragmentation'* function using the previous distances and then, we will extract the core area percentage and edge percentage statistics. Finally, we will plot the average of the patch core area percentage and edge percentage (% core area + % edge = 100%). 

```{r message = FALSE, warning = FALSE}
library(purrr)
Fragmentation_test.2 <- map_dfr(seq(100, 1000, 100), function(x){
  x.1 <- MK_Fragmentation(patches = vegetation_patches, edge_distance = x,
                          plot = FALSE)[[2]]
  CA <- mean(x.1$CAPercent)
  Edge <- mean(x.1$EdgePercent)
  x.2 <- rbind(data.frame('Edge distance' = x, Type = "Core Area", Percentage = CA),
                     data.frame('Edge distance' = x, Type = "Edge", Percentage = Edge))
  return(x.2)
})

head(Fragmentation_test.2)
```


```{r message = FALSE, warning = FALSE}
library(ggplot2)
ggplot(Fragmentation_test.2, aes(x = Edge.distance, y = Percentage, group = Type)) +
  geom_line(aes(color = Type))+
  geom_point(aes(color = Type))+ ylim(0,100)+
  scale_x_continuous("Edge depth distance (m)", labels = as.character(Fragmentation_test.2$Edge.distance), breaks = Fragmentation_test.2$Edge.distance)+
  scale_color_brewer(palette="Dark2")+
  theme_classic()
```


The average core area percentage (average patch area that has the least possible edge effect) for all patches decreases by more than 70% when considering an edge effect with an edge depth distance of 1 km.


 |Edge depth distance (m) | CoreArea (%)|
 |------------------------|:-----------:|
 |100                     | 83.5%       |
 |500                     | 34.14%      |
 |1000                    | 9.78%       |


<strong>Reference</strong>:

* INEGI. (2013). Conjunto de datos vectoriales de uso del suelo y vegetación, serie V (capa unión), escala 1:250,000. Instituto Nacional de Estadística y Geografía, Aguascalientes.

* McGarigal, K., S. A. Cushman, M. C. Neel, and E. Ene. 2002. FRAGSTATS: Spatial Pattern Analysis Program for Categorical Maps. Computer software program produced by the authors at the University of Massachusetts, Amherst. Available at the following web site: www.umass.edu/landeco/research/fragstats/fragstats.html.

* Haddad et al. (2015). Science Advances 1(2):e1500052. DOI: 10.1126/sciadv.1500052.
