---
title: "Protected Connected indicator"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Protected Connected indicator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", message=FALSE, warning=FALSE
)
```


```{r setup}
library(Makurhini)
library(raster)
```


## Overview
We will explore the ProtConn indicator which was developed to report international conservation targets (Saura et al. 2017), the indicator offers you an analysis of protected areas connectivity for a particular region.

## Loading data 
Loading two ecoregions and one set of protected areas, both shapefiles:

```{r polygons}
data("Protected_areas", package = "Makurhini")
data("regions", package = "Makurhini")
```

```{r echo=FALSE}
plot(regions, col=c("#FC8D62", "#8DA0CB", "#E78AC3"))
plot(Protected_areas, col = "#00B050", add=TRUE)
```

## ProtConn estimation for a single ecoregion
In the following example, we will estimate the ProtConn indicator and fractions in one ecoregion using two dispersal distances (10 and 30 km) and a connection probability of 0.5. Also, we will use a Transboundary buffer of 50 km (50000 meters) from the edge of the region (transboundary_type = "region", ?MK_ProtConn), the distance between protected areas will be using centroids.

```{r message=FALSE, warning=FALSE}
region <- regions[1,]
test.1 <- MK_ProtConn(nodes = Protected_areas, region = region, 
                      area_unit = "ha", distance = list(type= "centroid"), 
                      distance_thresholds = c(10000, 30000), probability = 0.5, 
                      transboundary = 50000, transboundary_type = "region",
                      LA = NULL, plot = TRUE, write = NULL,
                      intern = FALSE)
```

### Exploring the results for a single ecoregion

* Result 10 km:

```{r }
test.1$d10000$`Protected Connected (Viewer Panel)`
test.1$d10000
```


* Result 30 km:

```{r }
test.1$d30000$`Protected Connected (Viewer Panel)`
test.1$d30000
```

## ProtConn estimation for two or more ecoregions.

Now, we will use the three ecoregions. The processing time will be longer when using more regions, although we can reduce it using the parallel argument.

```{r message=FALSE, warning=FALSE}
test.2 <- MK_ProtConnMult(nodes = Protected_areas, regions = regions, 
                          area_unit = "ha", distance = list(type= "centroid"), 
                          distance_thresholds = c(10000, 30000), probability = 0.5, 
                          transboundary = 50000, transboundary_type = "region",
                          plot = FALSE, write = NULL, 
                          parallel = NULL, intern = FALSE)
```

### Exploring some results

* Table summary result:
```{r }
names(test.2)
test.2$ProtConn_10000$ProtConn_overall10000
```

* Shapefile result:
```{r }
test.2$ProtConn_10000$ProtConn_10000
```


```{r echo=FALSE}
ProtConn <- test.2[[1]]$ProtConn_10000
plot(ProtConn["ProtConn"])
```

* It is important not to forget that you can change the type of distance using the distance (<code>see,  distancefile() </code>) argument:

Euclidean distances:
* distance = list(type= "centroid")
* distance = list(type= "edge")

Least cost distances, you need a raster with resistance values, it is recommended that the range of values be from 1 to 10:
* distance = list(type= "least-cost", resistance = "resistance raster")
* distance = list(type= "commute-time", resistance = "resistance raster")


<strong>Reference</strong>:

* Saura, S., Bastin, L., Battistella, L., Mandrici, A., & Dubois, G. (2017). Protected areas in the world’s ecoregions: How well connected are they? Ecological Indicators, 76, 144–158. 
