---
title: "Protected Connected Land (ProtConn)"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{ProtConn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(Makurhini)
library(raster)
```

We will explore the ProtConn indicator which was developed to report international conservation targets (Saura et al. 2017), the indicator offers you an analysis of protected areas connectivity for a particular region.

* Loading data (two ecoregions and one set of protected areas, both shapefiles):

```{r message=FALSE, warning=FALSE}
#Protected areas
data("Protected_areas", package = "Makurhini")

#Ecoregions
data("regions", package = "Makurhini")

```


```{r message=FALSE, warning=FALSE, echo=FALSE}
plot(regions, col="#beaed4", main="Protected areas and ecoregions",
     axes = TRUE, key.pos = 4)
plot(Protected_areas, add=TRUE, col = "#00B050")
legend("topright",
       c("Ecoregions", "Protected areas"),
       fill = c("#beaed4", "#00B050"), # set the color of each legend line
       bty = "n", # turn off border
       cex = .8) # adjust font size
```

* ProtConn only one ecoregion using two distance thresholds (median distance) and a connection probability of 0.5. 

```{r message=FALSE, warning=FALSE}
region <- regions[1,]
test.1 <- MK_ProtConn(nodes = Protected_areas, region = region,
                    attribute = "Intersected area", area_unit = "ha",
                    distance = list(type= "centroid"),
                    distance_thresholds = c(50000, 10000),
                    probability = 0.5, transboundary = 50000,
                    LA = NULL, plot = TRUE, dPC = FALSE,
                    write = NULL, SAGA = FALSE, intern = FALSE)
```

Result 50 km:

```{r message=FALSE, warning=FALSE}
test.1$d50000
```


Result 10 km:

```{r message=FALSE, warning=FALSE}
test.1$d10000
```



* Using the two ecoregions (or ideally more regions):

```{r message=FALSE, warning=FALSE}
test.2 <- MK_ProtConnMult(nodes = Protected_areas, regions = regions,
                          attribute = "Intersected area", area_unit = "ha",
                          distance = list(type= "centroid"),
                          distance_thresholds = 10000,
                          probability = 0.5, transboundary = 50000,
                          plot = FALSE, write = NULL, 
                          intern = FALSE)
```

Table summary result:
```{r message=FALSE, warning=FALSE}
test.2$ProtConn_10000$ProtConn_10000

```

Shapefile result
```{r message=FALSE, warning=FALSE}
test.2$ProtConn_10000$ProtConn_overall_10000
```



Moreover, you can change distance using the distance (<code> distancefile () </code>) argument:

<code>Euclidean distances:
* distance = list(type= "centroid")
* distance = list(type= "edge")
Least cost distances:
* distance = list(type= "least-cost", resistance = "resistance raster")
* distance = list(type= "commute-time", resistance = "resistance raster")</code>


<strong>Reference</strong>:

* Saura, S., Bastin, L., Battistella, L., Mandrici, A., & Dubois, G. (2017). Protected areas in the world’s ecoregions: How well connected are they? Ecological Indicators, 76, 144–158. 
